---
title: "Connecting gene modules identified within different epilepsy cohorts"
author: "Liesbeth François"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
   html_document: 
     self_contained: yes
     fig_height: 6
     fig_width: 8
     keep_md: yes
     number_sections: yes
     theme: cerulean
     toc: yes
     toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=F, eval = T, include=F}
rm(list=ls())
gc() 
library(knitr)
library(kableExtra)
opts_chunk$set(
   include=FALSE,
   warning=FALSE, echo=FALSE, message=FALSE,
   concordance=TRUE
)
##
library(beeswarm)
library(BiocParallel)
library(DT)
library(edgeR)
library(ggplot2)
library(here)
library(limma)
library(MASS)
library(openxlsx)
library(plotly)
library(RColorBrewer)
library(readxl)
library(scales)
library(tibble)
library(tximport)
library(CRAFT)
library(reshape2)
library(WGCNA)
library(ggrepel)
library(UpSetR)
library(cluster)
library(gplots)
library(pheatmap)
library(CoReMo)
library(lme4)
library(ggpubr)
library(rlist)
library(dplyr)
library(BED)
library(heatmaply)
library(factoextra)
library(NbClust)
library(fgsea)
library(reactable)
library(stringr)
```

# Comparison of shared mechanisms across epilepsies

```{r}
load(here("Data/06_Modules/HS/Conservation/HS_ConservationResults.rda"))
all_cross <- crosstabulation 
## as the different cortex control samples are used for TSC, FCD2b and mTOR, they are not really independent results. We take the control assessment of mTOR, which includes all relevant controls for these conditions.
all_cons <- conservation %>% 
  filter(Condition %in% c("HS_disease","mTOR_disease", "TSC_disease", 
                          "FCD2b_disease", "FCD 2a", 
                          "HS_control", "mTOR_control")) 
##
load(here("Data/06_Modules/TSC/Conservation/TSC_ConservationResults.rda"))
all_cross <- crosstabulation %>% 
  bind_rows(all_cross)
all_cons <- conservation %>% 
  filter(Condition %in% c("HS_disease","mTOR_disease", "TSC_disease", 
                          "FCD2b_disease", "FCD 2a", 
                          "HS_control", "mTOR_control"))  %>% 
  bind_rows(all_cons)

##
load(here("Data/06_Modules/mTOR/Conservation/mTOR_ConservationResults.rda"))
all_cross <- crosstabulation %>% 
  bind_rows(all_cross)
all_cons <- conservation %>% 
  filter(Condition %in% c("HS_disease","mTOR_disease", "TSC_disease", 
                          "FCD2b_disease", "FCD 2a", 
                          "HS_control", "mTOR_control"))  %>% 
  bind_rows(all_cons)

##
load(here("Data/06_Modules/FCD2b/Conservation/FCD2b_ConservationResults.rda"))
all_cross <- crosstabulation %>% 
  bind_rows(all_cross)
all_cons <- conservation %>% 
  filter(Condition %in% c("HS_disease","mTOR_disease", "TSC_disease", 
                          "FCD2b_disease", "FCD 2a", 
                          "HS_control", "mTOR_control"))  %>% 
  bind_rows(all_cons)

all_cross1 <- all_cross %>% 
  mutate(m = paste(pmin(Module, `External module`), 
                   pmax(Module, `External module`), 
                   sep = "|")) %>% 
  filter(Project == "AMC") %>% 
  filter(Condition %in% c("HS_disease","mTOR_disease", "TSC_disease", 
                          "FCD2b_disease", "FCD 2a", 
                          "HS_control", "mTOR_control")) %>%  
  filter(!Module == `External module`) %>% 
  arrange(`P-Value`) 
```

```{r, include = TRUE, plotly = TRUE}
toPlot <- all_cons %>% 
  select(Module, Condition, p.value)
toPlot <- reshape2::dcast(toPlot, Condition~Module)
rownames(toPlot) <- toPlot[,1]
toPlot <- toPlot[,-1]
toPlot <-  data.matrix(toPlot)
toPlot[toPlot <= 0.0001] <- 0.0001
toPlot[toPlot > 0.05] <- 1
toPlot[is.na(toPlot)] <- 1
toPlot <- -log10(toPlot)
# dim(toPlot)
# Remove modules which are never conserved
toRem <- which(apply(toPlot, 2, function(x){sum(x)}) == 0)
toPlot <- toPlot[,-toRem]

heatmaply::heatmaply(toPlot, 
                     show_dend=c(TRUE, TRUE), 
                     hclust_method = "ward.D",
                     colors = viridis::viridis(n=256, 
                                               alpha = 1, 
                                               begin = 0, 
                                               end = 1,
                                               option = "cividis"),
                     k_col = 5,
                     fontsize_row = 5, 
                     fontsize_column = 5)  %>%
  layout(width = 1200)
```

# Identifying subclusters {.tabset}

## Clustering

```{r}
scope <- list(be = "Gene",
                source = "Ens_gene",
                organism = "human")
moduleList <- do.call(focusOnScope, c(
    list(
      x=getBekBeids(
        names = unique(all_cons$Module)),
      force=FALSE
    ),
    scope
  ))
names(moduleList) <- metadata(moduleList)$name
x <- lhclust(moduleList, distMeth = "inclusion", hcMeth = "ward.D2")
plot(x)

distMeth = "inclusion"
toPlot <- do.call(cbind, lapply(moduleList, function(z) {
        unlist(lapply(moduleList, function(y) {
            if (distMeth == "inclusion") {
                return(length(intersect(z, y))/min(length(z), 
                  length(y)))
            }
            if (distMeth == "jaccard") {
                return(length(intersect(z, y))/length(union(z, 
                  y)))
            }
        }))
    }))

```

```{r, include= TRUE}
p1 <- heatmaply(toPlot,
                 show_dend=c(TRUE, TRUE),
                 hclust_method = "ward.D2",
                show_dendrogram = c(FALSE,TRUE),
                 fontsize_row = 6,
                 fontsize_column = 6) 
p1
```

```{r, include=  TRUE}
# ## Elbow method
fviz_nbclust(toPlot, FUN = hcut,
             # hc_func = c("hclust"),
             # hc_method = "ward.D2", 
             method = "wss", 
             k.max = 100)
```

## Subclusters

```{r, include = TRUE}
## Targets
hc <- hclust(dist(1-toPlot), method = "ward.D2")
plot(hc)
subcluster <- cutree(hc, k = 29) ##52
rect.hclust(hc,k = 29, border=2:4)
```

```{r, include = TRUE}
stacked_subcluster <- stack(subcluster) %>% 
  mutate_if(is.factor, as.character) %>% 
  select(module = ind, group = values) %>% 
  arrange(group)
a <- stacked_subcluster %>% 
  inner_join(diffco_all, by = c("module" = "Module"))
print(stacked_subcluster)
```

```{r}
lstacked_subcluster <- split(x = stacked_subcluster, 
                            f = stacked_subcluster$group)

subcluster_enr <- lapply(lstacked_subcluster,
                      function(i){
  scope <- list(be = "Gene",
                source = "Ens_gene",
                organism = "human")
  ml <- do.call(focusOnScope, c(
    list(
      x=getBekBeids(
        names = pull(i,module)),
      force=FALSE
    ),
    scope
  ))
  names(ml) <- metadata(ml)$name
    
  nbCond <- length(unique(gsub("[.].*", "", names(ml))))
  
  toRet <- stack(ml) %>% 
    mutate_if(is.factor, as.character) %>% 
    as_tibble() %>% 
    add_count(values) %>% 
    filter(n >= nbCond & n > 1) %>%
    pull(values) %>%
    unique()
})
```


## Functional annotation {.tabset}


```{r}
bescope <- list(be = "Gene",
                source = "Ens_gene",
                organism = "human")
toTest <- rlist::list.flatten(subcluster_enr)
gn <- getBeIdDescription(stack(toTest)$values,  
                  be = "Gene", source = "Ens_gene", organism = "human")
eg <- unlist(do.call(focusOnScope, c(
    list(
      x=getBekBeids(analysisProjects = "AMC", type = "Background"),
      force=FALSE
    ),
    bescope
  )))
```

### MetaBase pathway maps

```{r}
faFile <- here("Data", "10_CohortComparison/FunctionalAnnotation-crosstabCluster_inclusionstringent.rda")
if(file.exists(faFile)){
  load(faFile)
}else{
  nbeidsLim <- c(10, 500)
  fl <- do.call(focusOnScope, c(
     list(
        x=getBekBeids(
           externalKnowledge="MetaBase_pathway",
           nbeids=nbeidsLim
        ),
        force=TRUE
     ),
     bescope
  ))
  tmp <- unlist(lapply(fl, length))
  fl <- fl[which(tmp>=min(nbeidsLim) & tmp <= max(nbeidsLim))]
  names(fl) <- metadata(fl)$externalID
  res <- TBTools::qrlEnrich(
     query=toTest, reference=fl,
     omega=eg,
     mc.cores=30
  )
  colnames(res) <- c(
     "uuid", "query", "ruuid", "reference",
     "intersection", "p-value", "FDR"
  )
  res <- do.call(rbind, by(
     res,
     res$uuid,
     function(d){
        d$rank <- rank(-log10(d$"p-value"))
        d$relRank <- d$rank/max(d$rank, na.rm=TRUE)
        return(d)
     }
  ))
  rownames(res) <- NULL
  resMB <- res %>% 
    mutate(type = "MetaBase_pathway") %>%
    inner_join(metadata(fl) %>% select("externalID", name), 
               by = c("ruuid" = "externalID")) %>%
    mutate(Pathway = paste0(sprintf('<a href="https://portal.genego.com/cgi/imagemap.cgi?id=%s" target="_blank">%s</a>',
                                    ruuid,
                                    name)))
  
    # Add list of genes intersect
  resMB$Intersect_genes <- apply(as.data.frame(resMB), 1, function(x){
      g <- intersect(toTest[[x["uuid"]]], fl[[x["ruuid"]]])
      paste(gn$symbol[match(g, gn$id)], collapse = ", ")
    })
}
```

```{r, include = FALSE}
datatable(resMB %>%
            filter(FDR <= 0.05) %>%
            select(Module = uuid,
                   `Module size` = query,
                   Concept = Pathway,
                   `Concept size` = reference,
                   Intersect = intersection,
                   FDR,
                   `Relative rank` = relRank) %>%
            mutate(Module = as.factor(Module)),
                   # Concept = as.factor(Concept)),
          filter = "top",
          escape = FALSE,
          options = list(
              order = list(5, "asc"),
              dom = "tp",
              pageLength = 6,
              order = list(list(3, "asc")),
              autoWidth = TRUE,
              columnDefs = list(list(width = '100px', targets = c(0, 2)))),
          rownames = FALSE) %>%
  formatSignif(6:7, digits = 2)
```

```{r, eval=TRUE, include = TRUE}
toShow <- resMB %>%
    filter(FDR <= 0.05) %>%
    mutate(Module = paste0("C", uuid)) %>%
    select(Module = Module,
           `Module size` = query,
           Concept = name,
           `Concept size` = reference,
           Intersect = intersection,
           FDR,
           `Relative rank` = relRank, 
           Intersect_genes, 
           ruuid) %>%
    mutate(Module = as.factor(Module),
           Concept = as.factor(Concept)) %>% 
  as.data.frame()
```

### Reactome pathways

```{r}
if(!file.exists(faFile)){
 nbeidsLim <- c(10, 500)
  fl <- do.call(focusOnScope, c(
     list(
        x=getBekBeids(
           externalKnowledge="Reactome_pathways_human",
           nbeids=nbeidsLim
        ),
        force=TRUE
     ),
     bescope
  ))
  tmp <- unlist(lapply(fl, length))
  fl <- fl[which(tmp>=min(nbeidsLim) & tmp <= max(nbeidsLim))]
  names(fl) <- metadata(fl)$externalID
  res <- qrlEnrich(
     query=toTest, reference=fl,
     omega=eg,
     mc.cores=30
  )
  colnames(res) <- c(
     "uuid", "query", "ruuid", "reference",
     "intersection", "p-value", "FDR"
  )
  res <- do.call(rbind, by(
     res,
     res$uuid,
     function(d){
        d$rank <- rank(-log10(d$"p-value"))
        d$relRank <- d$rank/max(d$rank, na.rm=TRUE)
        return(d)
     }
  ))
  rownames(res) <- NULL
  resRA <- res %>% 
    mutate(type = "Reactome_pathways_human") %>%
    inner_join(metadata(fl) %>% select(externalID, name), 
               by = c("ruuid" = "externalID")) %>%
    mutate(Pathway = paste0(sprintf('<a href="http://www.reactome.org/PathwayBrowser/#/%s" target="_blank">%s</a>',
                                    ruuid,
                                    name)))
  
      # Add list of genes intersect
  resRA$Intersect_genes <- apply(as.data.frame(resRA), 1, function(x){
      g <- intersect(toTest[[x["uuid"]]], fl[[x["ruuid"]]])
      paste(gn$symbol[match(g, gn$id)], collapse = ", ")
    })
}
```

```{r, include = FALSE}
datatable(resRA %>%
            filter(FDR <= 0.05) %>%
            select(Module = uuid,
                   `Module size` = query,
                   Concept = Pathway,
                   `Concept size` = reference,
                   Intersect = intersection,
                   FDR,
                   `Relative rank` = relRank)%>%
            mutate(Module = as.factor(Module)),
                   # Concept = as.factor(Concept)),
          filter = "top",
          escape = FALSE,
          options = list(
              order = list(5, "asc"),
              dom = "tp",
              pageLength = 6,
              order = list(list(3, "asc")),
              autoWidth = TRUE,
              columnDefs = list(list(width = '100px', targets = c(0, 2)))),
          rownames = FALSE) %>%
  formatSignif(6:7, digits = 2)
```

```{r, eval=TRUE, include = TRUE}
toShow <- resRA %>%
    filter(FDR <= 0.05) %>%
    mutate(Module = paste0("C", uuid)) %>%
    select(Module = Module,
           `Module size` = query,
           Concept = name,
           `Concept size` = reference,
           Intersect = intersection,
           FDR,
           `Relative rank` = relRank, 
           Intersect_genes, 
           ruuid) %>%
    mutate(Module = as.factor(Module),
           Concept = as.factor(Concept)) %>% 
  as.data.frame()
```

### Gene Ontology: biological process

```{r}
if(!file.exists(faFile)){
  nbeidsLim <- c(10, 500)
  fl <- do.call(focusOnScope, c(
     list(
        x=getBekBeids(
           externalKnowledge="GO_biological_process_human",
           nbeids=nbeidsLim
        ),
        force=TRUE
     ),
     bescope
  ))
  tmp <- unlist(lapply(fl, length))
  fl <- fl[which(tmp>=min(nbeidsLim) & tmp <= max(nbeidsLim))]
  names(fl) <- metadata(fl)$externalID
  res <- qrlEnrich(
     query=toTest, reference=fl,
     omega=eg,
     mc.cores=30
  )
  colnames(res) <- c(
     "uuid", "query", "ruuid", "reference",
     "intersection", "p-value", "FDR"
  )
  res <- do.call(rbind, by(
     res,
     res$uuid,
     function(d){
        d$rank <- rank(-log10(d$"p-value"))
        d$relRank <- d$rank/max(d$rank, na.rm=TRUE)
        return(d)
     }
  ))
  rownames(res) <- NULL
  resGOBP <- res %>% 
    mutate(type = "GO_biological_process_human") %>%
    inner_join(metadata(fl) %>% select(externalID, name), 
               by = c("ruuid" = "externalID")) %>%
    mutate(Pathway = paste0(sprintf('<a href="http://amigo.geneontology.org/amigo/term/GO:%s" target="_blank">%s</a>',
                                    ruuid,
                                    name)))
  
      # Add list of genes intersect
  resGOBP$Intersect_genes <- apply(as.data.frame(resGOBP), 1, function(x){
      g <- intersect(toTest[[x["uuid"]]], fl[[x["ruuid"]]])
      paste(gn$symbol[match(g, gn$id)], collapse = ", ")
    })
}
```

```{r, include = FALSE}
datatable(resGOBP %>%             
            filter(FDR <= 0.05) %>%
            select(Module = uuid,
                   `Module size` = query,
                   Concept = Pathway,
                   `Concept size` = reference,
                   Intersect = intersection,
                   FDR,
                   `Relative rank` = relRank) %>%
            mutate(Module = as.factor(Module)),
                   # Concept = as.factor(Concept)),
          filter = "top",
          escape = FALSE,
          options = list(
              order = list(5, "asc"),
              dom = "tp",
              pageLength = 6,
              order = list(list(3, "asc")),
              autoWidth = TRUE,
              columnDefs = list(list(width = '100px', targets = c(0, 2)))),
          rownames = FALSE) %>%
  formatSignif(6:7, digits = 2)
```

```{r, eval=TRUE, include = TRUE}
toShow <- resGOBP %>%
    filter(FDR <= 0.05) %>%
    mutate(Module = paste0("C", uuid)) %>%
    select(Module= Module,
           `Module size` = query,
           Concept = name,
           `Concept size` = reference,
           Intersect = intersection,
           FDR,
           `Relative rank` = relRank, 
           Intersect_genes, 
           ruuid) %>%
    mutate(Module = as.factor(Module),
           Concept = as.factor(Concept)) %>% 
  as.data.frame()
```

### Gene Ontology: molecular function

```{r}
if(!file.exists(faFile)){

  nbeidsLim <- c(10, 500)
  fl <- do.call(focusOnScope, c(
     list(
        x=getBekBeids(
           externalKnowledge="GO_molecular_function_human",
           nbeids=nbeidsLim
        ),
        force=TRUE
     ),
     bescope
  ))
  tmp <- unlist(lapply(fl, length))
  fl <- fl[which(tmp>=min(nbeidsLim) & tmp <= max(nbeidsLim))]
  names(fl) <- metadata(fl)$externalID
  res <- qrlEnrich(
     query=toTest, reference=fl,
     omega=eg,
     mc.cores=30
  )
  colnames(res) <- c(
     "uuid", "query", "ruuid", "reference",
     "intersection", "p-value", "FDR"
  )
  res <- do.call(rbind, by(
     res,
     res$uuid,
     function(d){
        d$rank <- rank(-log10(d$"p-value"))
        d$relRank <- d$rank/max(d$rank, na.rm=TRUE)
        return(d)
     }
  ))
  rownames(res) <- NULL
  resGOMF <- res %>% 
    mutate(type = "GO_molecular_function_human") %>%
    inner_join(metadata(fl) %>% select(externalID, name), 
               by = c("ruuid" = "externalID")) %>%
    mutate(Pathway = paste0(sprintf('<a href="http://amigo.geneontology.org/amigo/term/GO:%s" target="_blank">%s</a>',
                                    ruuid,
                                    name)))
  
      # Add list of genes intersect
  resGOMF$Intersect_genes <- apply(as.data.frame(resGOMF), 1, function(x){
      g <- intersect(toTest[[x["uuid"]]], fl[[x["ruuid"]]])
      paste(gn$symbol[match(g, gn$id)], collapse = ", ")
    })
}
```

```{r, include = FALSE}
datatable(resGOMF %>%          
            filter(FDR <= 0.05) %>%
            select(Module = uuid,
                   `Module size` = query,
                   Concept = Pathway,
                   `Concept size` = reference,
                   Intersect = intersection,
                   FDR,
                   `Relative rank` = relRank) %>%
            mutate(Module = as.factor(Module)),
                   # Concept = as.factor(Concept)),
          filter = "top",
          escape = FALSE,
          options = list(
              order = list(5, "asc"),
              dom = "tp",
              pageLength = 6,
              order = list(list(3, "asc")),
              autoWidth = TRUE,
              columnDefs = list(list(width = '100px', targets = c(0, 2)))),
          rownames = FALSE) %>%
  formatSignif(6:7, digits = 2)
```

```{r, eval=TRUE, include = TRUE}
toShow <- resGOMF %>%
    filter(FDR <= 0.05) %>%
    mutate(Module = paste0("C", uuid)) %>%
    select(Module = Module,
           `Module size` = query,
           Concept = name,
           `Concept size` = reference,
           Intersect = intersection,
           FDR,
           `Relative rank` = relRank, 
           Intersect_genes, 
           ruuid) %>%
    mutate(Module = as.factor(Module),
           Concept = as.factor(Concept)) %>% 
  as.data.frame()

```

### Gene Ontology: cellular component

```{r}
if(!file.exists(faFile)){

  nbeidsLim <- c(10, 500)
  fl <- do.call(focusOnScope, c(
     list(
        x=getBekBeids(
           externalKnowledge="GO_cellular_component_human",
           nbeids=nbeidsLim
        ),
        force=TRUE
     ),
     bescope
  ))
  tmp <- unlist(lapply(fl, length))
  fl <- fl[which(tmp>=min(nbeidsLim) & tmp <= max(nbeidsLim))]
  names(fl) <- metadata(fl)$externalID
  res <- qrlEnrich(
     query=toTest, reference=fl,
     omega=eg,
     mc.cores=30
  )
  colnames(res) <- c(
     "uuid", "query", "ruuid", "reference",
     "intersection", "p-value", "FDR"
  )
  res <- do.call(rbind, by(
     res,
     res$uuid,
     function(d){
        d$rank <- rank(-log10(d$"p-value"))
        d$relRank <- d$rank/max(d$rank, na.rm=TRUE)
        return(d)
     }
  ))
  rownames(res) <- NULL
  resGOCC <- res %>% 
    mutate(type = "GO_cellular_component_human") %>%
    inner_join(metadata(fl) %>% select(externalID, name), 
               by = c("ruuid" = "externalID")) %>%
    mutate(Pathway = paste0(sprintf('<a href="http://amigo.geneontology.org/amigo/term/GO:%s" target="_blank">%s</a>',
                                    ruuid,
                                    name)))
  
      # Add list of genes intersect
  resGOCC$Intersect_genes <- apply(as.data.frame(resGOCC), 1, function(x){
      g <- intersect(toTest[[x["uuid"]]], fl[[x["ruuid"]]])
      paste(gn$symbol[match(g, gn$id)], collapse = ", ")
    })
}
```

```{r, include = FALSE}
datatable(resGOCC %>%              
            filter(FDR <= 0.05) %>%
            select(Module = uuid,
                   `Module size` = query,
                   Concept = Pathway,
                   `Concept size` = reference,
                   Intersect = intersection,
                   FDR,
                   `Relative rank` = relRank) %>%
            mutate(Module = as.factor(Module)),
                   # Concept = as.factor(Concept)),
          filter = "top",
          escape = FALSE,
          options = list(
              order = list(5, "asc"),
              dom = "tp",
              pageLength = 6,
              order = list(list(3, "asc")),
              autoWidth = TRUE,
              columnDefs = list(list(width = '100px', targets = c(0, 2)))),
          rownames = FALSE) %>%
  formatSignif(6:7, digits = 2)
```

```{r, eval=TRUE, include = TRUE}
toShow <- resGOCC %>%
    filter(FDR <= 0.05) %>%
    mutate(Module = paste0("C", uuid)) %>%
    select(Module = Module,
           `Module size` = query,
           Concept = name,
           `Concept size` = reference,
           Intersect = intersection,
           FDR,
           `Relative rank` = relRank, 
           Intersect_genes, 
           ruuid) %>%
    mutate(Module = as.factor(Module),
           Concept = as.factor(Concept)) %>% 
  as.data.frame()
```

### PangloaDB cell type biomarkers

```{r}
if(!file.exists(faFile)){
  tkcon <- chTKCat(password = "")
  pangloadb <- get_MDB(tkcon,"PanglaoDB")
  fl <- inner_join(pangloadb$PanglaoDB_Markers,
                   pangloadb$PanglaoDB_Genes %>% select(symbol, ensembl),
                   by = c("symbol")) %>% 
    inner_join(pangloadb$PanglaoDB_CellTypes %>% select(`cell type`, organ),
               by = c("cell type")) %>% 
    filter(grepl("ENSG", ensembl))
  sfl <- split(fl$ensembl, fl$`cell type`)
  
  res <- qrlEnrich(
     query=toTest, reference=sfl,
     omega=unlist(eg),
     mc.cores=30
  )
  colnames(res) <- c(
     "uuid", "query", "ruuid", "reference",
     "intersection", "p-value", "FDR"
  )
  res <- do.call(rbind, by(
     res,
     res$uuid,
     function(d){
        d$rank <- rank(-log10(d$"p-value"))
        d$relRank <- d$rank/max(d$rank, na.rm=TRUE)
        return(d)
     }
  ))
  rownames(res) <- NULL
  resPAN <- res %>% 
    mutate(type = "PangloaDB cell type biomarkers") %>% 
    mutate(Organ = fl$organ[match(ruuid, fl$`cell type`)])
  
      # Add list of genes intersect
  resPAN$Intersect_genes <- apply(as.data.frame(resPAN), 1, function(x){
      g <- intersect(toTest[[x["uuid"]]], fl[[x["ruuid"]]])
      paste(gn$symbol[match(g, gn$id)], collapse = ", ")
    })
  
  
  save(resMB, resRA, resGOBP, resGOMF, resGOCC, resHB, resPAN,
     file = faFile)
}
```

```{r, include = FALSE}
datatable(resPAN %>%        
            filter(FDR <= 0.05) %>%
            select(Module = uuid,
                   `Module size` = query,
                   Concept = ruuid,
                   Organ,
                   `Concept size` = reference,
                   Intersect = intersection,
                   FDR,
                   `Relative rank` = relRank) %>%
            mutate(Module = as.factor(Module)),
                   # Concept = as.factor(Concept)),
          filter = "top",
          options = list(
              order = list(5, "asc"),
              dom = "tp",
              pageLength = 6,
              order = list(list(3, "asc")),
              autoWidth = TRUE,
              columnDefs = list(list(width = '100px', targets = c(0, 2)))),
          rownames = FALSE) %>%
  formatSignif(6:7, digits = 2)
```

```{r, eval=TRUE, include = TRUE}
toShow <- resPAN %>%
    filter(FDR <= 0.05) %>%
  filter(Organ == "Brain") %>% 
    mutate(Module = paste0("C", uuid)) %>%
    select(Module = Module,
           `Module size` = query,
           Concept = ruuid,
           Organ,
           `Concept size` = reference,
           Intersect = intersection,
           FDR,
           `Relative rank` = relRank, 
           Intersect_genes) %>%
    mutate(Module = as.factor(Module),
           Concept = as.factor(Concept)) %>% 
  as.data.frame()


reactable(toShow %>% 
            mutate_if(is.numeric, signif, 2),
          filterable=TRUE,
          resizable=TRUE, 
          wrap=FALSE,
          fullWidth=TRUE,
          defaultSorted = c("FDR"),
          columns = list(
             Module=colDef(width=100),
             Concept = colDef(width=450)
            ),
           details = function(index){
             x <- toShow[index,] %>% 
               pull(Intersect_genes) %>% 
               str_split(", ") %>% 
               unlist()
             toShow2 <- tibble(Gene = x, 
                              Ens_gene = gn$id[match(x, gn$symbol)])
             htmltools::div(
                reactable(toShow2, 
                          columns = list(
                             Ens_gene = colDef(cell = function(value, index) {
                                  # Render as a link
                                  url <- sprintf("http://www.ensembl.org/id/%s", 
                                                 value)
                                  htmltools::tags$a(href = url, target = "_blank", as.character(value))
                                }))
                          )
                )
           }
          )

```

## Gene content

```{r}
ann <- bind_rows(resMB, resPAN, resRA, resGOBP, resGOMF) %>% 
  rename(module = uuid)

pathway_preference <- c("PangloaDB cell type biomarkers", "Reactome_pathways_human",
                        "MetaBase_pathway", "GO_biological_process_human",
                        "GO_molecular_function_human" )
pathway_preference <- factor(pathway_preference,
                        levels = c("PangloaDB cell type biomarkers",
                                   "Reactome_pathways_human",
                        "MetaBase_pathway", "GO_biological_process_human",
                        "GO_molecular_function_human" ))


toRet1 <- ann %>% dplyr::group_by(module) %>% 
  dplyr::filter(FDR <= 0.05) %>% dplyr::top_n(n = 1, 
  wt = relRank) %>% dplyr::ungroup() %>% dplyr::select(module = module, 
  concept = name, externalID = ruuid, type, rank, 
  relRank, `p-value`, FDR, modSize = query, conceptSize = reference, intersection)
toRet1 <- toRet1 %>% 
  dplyr::filter(type %in% pathway_preference) %>% 
  dplyr::mutate(type = factor(type, level = levels(pathway_preference))) %>% 
  dplyr::arrange(type) %>% 
  dplyr::distinct(module, .keep_all = TRUE) %>% 
  dplyr::mutate(type = as.character(type))


annotation_cluster <- toRet1 %>% 
  dplyr::mutate(geneId = NA_character_, geneSymbol = NA_character_, 
                geneName = NA_character_) %>% 
  dplyr::select(module, geneId, geneSymbol, geneName, 
                pathwayId = externalID, pathwayName = concept, 
                pathwayResource = type) %>% 
  mutate(pathway = case_when(is.na(pathwayName) ~ pathwayId,
                             TRUE ~ pathwayName))

toShow <- stack(subcluster_enr) %>% 
  as_tibble()
toShow <- inner_join(toShow,
                     annotation_cluster,
                     by = c("ind" = "module")) %>% 
  inner_join(getBeIdDescription(ids = .$values, be = "Gene", source = "Ens_gene", 
                                organism = "human"),
             by = c("values" = "id")) %>% 
  mutate(Gene = paste0(sprintf('<a href="%s" target="_blank">%s</a>',
                                    getBeIdURL(values, 
                                               databases = "Ens_gene"),
                                    symbol)))
```

```{r, include = TRUE}
datatable(toShow %>% select(Cluster = ind, 
                            Gene, name, 
                            pathway,   
                            pathwayResource),
          escape = FALSE,
          rownames = FALSE,
          filter = 'top',
          fillContainer = FALSE)
```

## Common TF per cluster {.tabset}

```{r, include = TRUE, results='asis', plotly = TRUE, fig.width = 10, fig.height = 15, echo = FALSE}
hpa <- get_MDB(tkcon, "HPA")

modules_data_plot <- list()

for(i in 1:max(stacked_subcluster$group)){
  toSel <- (stacked_subcluster %>% filter(group == i) %>% pull(module))
  x <- all_cross1 %>%
        filter(Module %in%  toSel & 
                 `External module` %in% toSel) %>%
        arrange(`P-Value`) %>%
        filter(`P-Value` <= 0.05) %>% 
        filter(Module != `External module`) %>%
        distinct()
  common_regulators <- all_modules %>% 
    filter(uuid %in% toSel) %>% 
    arrange(`p-value`) %>% 
    distinct(uuid, value, .keep_all = T) %>% 
    add_count(value) 
  reg <- common_regulators %>% 
    filter(n == max(n) & n > 1) %>% 
    mutate(col = case_when(impact == "Activation" ~ "#1a9850",
                       TRUE ~ "#d73027")) %>% 
    mutate(gene = getBeIdDescription(ids = value, 
                                     be = "Object", 
                                     source = "MetaBase_object", 
                                     organism = "human")$symbol)

  col <- c("TSC" = "#d95f02",
             "mTOR" = "#7570b3",
             "FCD2b" = "#e7298a", 
             "HS" = "#66a61e")

  
  nodes <- all_cross %>% 
    filter(Module %in% toSel) %>% 
    select(id = Module, module = `Module size`, label = Module, Condition)  %>% 
    mutate(Condition = gsub("[.].*", "", id)) %>%
    arrange(desc(module)) %>% 
    distinct(id, .keep_all = TRUE) %>%
    mutate(color = col[Condition],
           group = "Module") %>% 
    bind_rows(tibble(id = unique(pull(reg, gene)),
                     module = NA,
                     label = id,
                     Condition = NA_character_,
                     group = "Regulator",
                     color = "#b3cde3"))  %>% 
    distinct()
  edges <- x %>% 
    filter(intersect != 0) %>%
    mutate(label = intersect) %>%
    mutate(perc = round(intersect*2/(`Module size` + `External size`), 2)) %>% 
    mutate(m = paste(pmin(Module, `External module`), pmax(Module, `External module`),
                     sep = "_"),
           color = "#878787") %>% 
    distinct(m, .keep_all = T) %>% 
    select(from = Module, to = `External module`, label = perc, title = intersect, color) %>% 
    distinct() %>% 
    bind_rows(tibble(from = reg$gene,
                     to = reg$uuid,
                     label = reg$intersection,
                     title = reg$intersection,
                     color = reg$col,
                     arrows = "to"))  %>% 
    distinct()
  
  modules_data_plot[[i]] <- list(nodes = nodes,
                                 edges = edges,
                                 gene_overlap_modules = x,
                                 common_regulators = common_regulators) 
  
  cat(sprintf("\n\n### Cluster %s\n\n", i))
  cat("\n\n")
  print(htmltools::tagList(
    visNetwork(nodes = nodes, 
             edges = edges, randomSeed =111) %>% 
      visPhysics(solver = "forceAtlas2Based", 
                 forceAtlas2Based = list(gravitationalConstant = -50)) %>% 
        # darkblue square with shadow for group "A"
      visGroups(groupname = "Module", shape = "round", 
                shadow = list(enabled = TRUE)) %>% 
      # red triangle for group "B"
      visGroups(groupname = "Regulator", shape = "triangle") 
  ))
  cat("\n\n")
}
names(modules_data_plot) <- sort(unique(stacked_subcluster$group))
save(modules_data_plot, 
       file = here("Data/14_Main_Manuscript_Figures/shared_modules_network_data.rda"))
```

```{r}
visNetwork(nodes = nodes, 
             edges = edges, randomSeed =111) %>% 
  visPhysics(solver = "forceAtlas2Based", 
             forceAtlas2Based = list(gravitationalConstant = -50)) %>% 
    # darkblue square with shadow for group "A"
  visGroups(groupname = "Module", shape = "round", 
            shadow = list(enabled = TRUE)) %>% 
  # red triangle for group "B"
  visGroups(groupname = "Regulator", shape = "triangle")   
```

## Conservation of regulation in other epilepsies and healthy tissue

```{r, include = TRUE}
toShow <- all_cons %>% 
  select(-Project) %>% 
  mutate(Cluster = stacked_subcluster$group[match(Module, 
                                                  stacked_subcluster$module)]) %>% 
  arrange(Cluster) %>% 
  select(-R2, -p.value) %>% 
  tidyr::pivot_wider(names_from = Condition, values_from = FDR)
```

```{r}
mb <- resMB %>% 
  group_by(uuid) %>% 
  filter(FDR <= 0.05) %>% 
  arrange(FDR) %>% 
  slice(1:2) %>% 
  summarize(pw = paste(name, collapse = " | \n")) %>% 
  ungroup()
ra <- resRA %>% 
  group_by(uuid) %>% 
  filter(FDR <= 0.05) %>% 
  arrange(FDR) %>% 
  slice(1:2) %>% 
  summarize(pw = paste(name, collapse = " | \n")) %>% 
  ungroup()
pan <- resPAN %>% 
  group_by(uuid) %>% 
  filter(FDR <= 0.05) %>% 
  arrange(FDR) %>% 
  slice(1:2) %>% 
  summarize(pw = paste(ruuid, collapse = " | \n")) %>% 
  ungroup()
gobp <- resGOBP %>% 
  group_by(uuid) %>% 
  filter(FDR <= 0.05) %>% 
  arrange(FDR) %>% 
  slice(1:2) %>% 
  summarize(pw = paste(name, collapse = " | \n")) %>% 
  ungroup()
gomf <- resGOMF %>% 
  group_by(uuid) %>% 
  filter(FDR <= 0.05) %>% 
  arrange(FDR) %>% 
  slice(1:2) %>% 
  summarize(pw = paste(name, collapse = " | \n")) %>% 
  ungroup()

toWrite <- toShow %>% 
  left_join(diffco_all,
             by = c("Module")) %>% 
  rename("FCD2a_disease" = "FCD 2a",
         Cortex_control = mTOR_control, 
         Hippocampus_control = HS_control) %>% 
  mutate(MetaBase = mb$pw[match(Cluster, mb$uuid)],
         Reactome = ra$pw[match(Cluster, ra$uuid)],
         PangloaDB = pan$pw[match(Cluster, pan$uuid)],
         GO_BiologicalProcess = gobp$pw[match(Cluster, gobp$uuid)],
         GO_MolecularFunction = gomf$pw[match(Cluster, gomf$uuid)]) %>% 
  select(Module, Cluster,
         14:18,
         10:13,
         3:9)
write.csv(toWrite, 
          file = here("Data/10_CohortComparison/summary_cohort_comparison.csv"))
```

