---
title: "Discriminant analysis on different epilepsy and healthy cohorts"
author: "Liesbeth François"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
   html_document: 
     self_contained: yes
     fig_height: 6
     fig_width: 8
     keep_md: yes
     number_sections: yes
     theme: cerulean
     toc: yes
     toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=F, eval = T, include=F}
rm(list=ls())
gc() 
library(knitr)
library(kableExtra)
opts_chunk$set(
   include=FALSE,
   warning=FALSE, echo=FALSE, message=FALSE,
   concordance=TRUE
)
##
library(beeswarm)
library(DT)
library(edgeR)
library(ggplot2)
library(here)
library(limma)
library(MASS)
library(openxlsx)
library(plotly)
library(RColorBrewer)
library(readxl)
library(tibble)
library(reshape2)
library(ggrepel)
library(ggpubr)
library(heatmaply)
library(factoextra)
library(NbClust)
library(adegenet)
library(dplyr)
library(ggdendro)

tkcon <- chTKCat(user = "default", password = "")
pdb <- chTKCat(host="bel038783", port=9311L, user="lfrancois", password=cred) %>%
  get_MDB("AMC_collaboration", check = FALSE) ## internal database which stores the output of the different coremo analyses for ease of access
```

```{r}
## Load sample information from all cohorts
amc_subjects <- pdb$AMC_subjects 
amc_samples <- pdb$AMC_samples %>% 
  inner_join(amc_subjects, by = c("subject")) %>% 
  filter(subject %in% amc_subjects$subject) %>% 
  mutate(disease = case_when(disease == "Control" ~ paste(disease, `tissue generic origin`, sep = "_"),
                             TRUE ~ disease)) %>% 
  mutate(disease = gsub(paste("Frontal", "Parietal", "Temporal", sep = "|"), "Cortex", disease)) %>% 
  mutate(Cohort = case_when(disease == "Focal cortical dysplasia type 2a" ~ "FCD IIa",
                          disease == "Focal cortical dysplasia type 2b" ~ "FCD IIb",
                          disease == "Tuberous Sclerosis Complex" ~ "TSC",
                          disease == "Hippocampal Sclerosis" ~ "TLE+HS",
                          disease == "Control_Cortex" ~ "Cortex",
                          TRUE ~ "Hippocampus"))
table(amc_samples$Cohort)

d <-  pdb$AMC_rnaseq[,amc_samples$sample]
dim(d)
dim(amc_samples)
```

# Clustering and dendrogram

```{r,eval=T,include=T,echo=F}
pca <- prcomp(t(d), center=TRUE, scale.=TRUE)
hc <- hclust(dist(1-pca$x), method = "ward.D2")

dendrogram_data <- dendro_data(as.dendrogram(hc))
dendrogram_segments <- dendrogram_data$segments # contains all dendrogram segment data
dendrogram_ends <- dendrogram_segments %>%
 filter(yend == 0) %>% # filter for terminal dendrogram ends
 left_join(dendrogram_data$labels, by = "x") %>% 
 rename(sample_name = label) %>%
 left_join(toPlot, by = c("sample_name" = "sample")) %>% 
  mutate(Cohort = case_when(Cohort == "TLE+HS" ~ "TLE",
                            TRUE ~ Cohort))


unique_vars <- levels(factor(dendrogram_ends$disease)) %>% 
 as.data.frame() %>% rownames_to_column("row_id") 
# count number of unique variables
color_count <- length(unique(unique_vars$.))
# get RColorBrewer palette
get_palette <- colorRampPalette(brewer.pal(n = 8, name = "Set1"))
# produce RColorBrewer palette based on number of unique variables in metadata:
palette <- get_palette(color_count) %>% 
 as.data.frame() %>%
 rename("color" = ".") %>%
 rownames_to_column(var = "row_id")
color_list <- left_join(unique_vars, palette, by = "row_id") %>%
 select(-row_id)
species_color <- as.character(color_list$color)
names(species_color) <- color_list$.

cols <- control_palettes(palette = 'Categorical_12_colors')[c(12, 1, 2, 5, 4, 3),] %>% 
  pull(hex)

p1 <- ggplot() +
 geom_segment(data = dendrogram_segments, 
              aes(x=x, y=y, xend=xend, yend=yend)) +
 geom_segment(data = dendrogram_ends,
              aes(x=x, y=y.x, xend=xend, yend=yend, 
                  color = Cohort, 
                  label = paste('sample name: ', sample_name, 
                                '<br>', 'species: ', disease))
              ) + 
 scale_color_manual(values = cols) +
 scale_y_reverse() +
 coord_flip() + 
 theme_void() + #theme(legend.position = 'none') + 
 ylab('Distance') +
 xlab("") + 
 theme(axis.text=element_blank(), legend.title = element_text(size = 18), 
       legend.text = element_text(size = 16)) 
p1
```

# Discriminant analysis of principal components (DAPC)

In this multivariate statistical approach variance in the sample is partitioned into a between-group and within- group component, in an effort to maximize discrimination between groups. In DAPC, data is first transformed using a principal components analysis (PCA) and subsequently clusters are identified using discriminant analysis (DA). Here DAPC is used in a supervised manner, where the groups (disease cohorts) are provided to perform the discriminant analysis.

More information from [DAPC tutorial](https://grunwaldlab.github.io/Population_Genetics_in_R/DAPC.html)

## All cohorts

```{r, include = T}
grp <- amc_samples %>% 
      mutate(disease = case_when(disease == "Focal cortical dysplasia type 2a" ~ "FCD IIa",
                          disease == "Focal cortical dysplasia type 2b" ~ "FCD IIb",
                          disease == "Tuberous Sclerosis Complex" ~ "TSC",
                          disease == "Hippocampal Sclerosis" ~ "TLE-HS",
                          disease == "Control_Cortex" ~ "Cortex",
                          TRUE ~ "Hippocampus")) %>% 
    pull(disease) %>% as.factor() %>% as.numeric()
names(grp) <- pull(amc_samples, sample)

grp_name <- amc_samples %>% 
      mutate(disease = case_when(disease == "Focal cortical dysplasia type 2a" ~ "FCD IIa",
                          disease == "Focal cortical dysplasia type 2b" ~ "FCD IIb",
                          disease == "Tuberous Sclerosis Complex" ~ "TSC",
                          disease == "Hippocampal Sclerosis" ~ "TLE-HS",
                          disease == "Control_Cortex" ~ "Cortex",
                          TRUE ~ "Hippocampus")) %>% 
    pull(disease) %>% as.factor() 
grp_name <- tibble(name = levels(grp_name),
                   value = 1:6)

if(file.exists(here("Data/12_PLSDA_DAPC/dapc_disease_clusters"))){
  load(here("Data/12_PLSDA_DAPC/dapc_disease_clusters"))
}else{
 dapc <- dapc(t(d), grp = grp, n.pca = 7, n.da = 5)
 save(dapc, file = "/home/lfrancois/Documents/projects/AMC_collaboration_2019/Data/12_PLSDA_DAPC/dapc_disease_clusters")
}

# tmp <- optim.a.score(dapc2)
summary(dapc2)

assigment <- tibble(prior = dapc$grp,
                     posterior = dapc$assign) %>% 
  mutate(prior_name = grp_name$name[match(prior, grp_name$value)],
         post_name = grp_name$name[match(posterior,
                                         grp_name$value)])
h <- table(assigment$prior_name, assigment$post_name)
h <- melt(h) %>% 
  as_tibble() %>% 
  mutate(Var2 = factor(Var2,
                       levels = c("Hippocampus", "TLE-HS", 
                                  "Cortex", "FCD IIa", "FCD IIb", "TSC")),
         Var1 = factor(Var1,
                       levels = c("Hippocampus", "TLE-HS", 
                                  "Cortex", "FCD IIa", "FCD IIb", "TSC")))
p2 <- ggplot(h, aes(x = Var1, y = Var2, fill = log10(value), label = value)) +
  geom_tile( color = "grey") +
  geom_text() +
  scale_fill_gradient2(na.value = "white") +
  labs(x = "Prior assignment", y = "Posterior assignment") + 
  theme_minimal() +
  theme(legend.pos = "none", text = element_text(size=14),
        legend.text = element_text(size = 16),
        axis.text=element_text(size=20), axis.title = element_text(size=18)) 
p2
ggsave(here("Data/12_PLSDA_DAPC/reassignment_all_cohorts.png"))


toPlot <- dapc$ind.coord %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "sample") %>% 
  inner_join(amc_samples %>% select(sample, disease, 
                                    `tissue generic origin`, age),
             by = c("sample")) %>% 
  mutate(grp = dapc$grp[sample]) %>% 
  as_tibble() %>% 
  mutate(Cohort = case_when(disease == "Focal cortical dysplasia type 2a" ~ "FCD IIa",
                          disease == "Focal cortical dysplasia type 2b" ~ "FCD IIb",
                          disease == "Tuberous Sclerosis Complex" ~ "TSC",
                          disease == "Hippocampal Sclerosis" ~ "TLE-HS",
                          disease == "Control_Cortex" ~ "Cortex",
                          TRUE ~ "Hippocampus"))
```

```{r}
cols <- control_palettes(palette = 'Categorical_12_colors')[c(12, 1, 2, 5, 4, 3),] %>% 
  pull(hex)

p3 <- ggplot(toPlot, aes(x = LD1, y = LD2, col = Cohort)) + 
  geom_point(size = 3, alpha = 0.7) +
  stat_ellipse(level=0.95) +
  # scale_color_brewer(palette = "Set2") +
  scale_color_manual(values=cols) +
  theme_bw() +
  theme(legend.pos = "none") +
  xlab("LD1") +
  ylab("LD2") +
  theme(legend.pos = "none", text = element_text(size = 16), 
        axis.title = element_text(size = 18), axis.text=element_text(size=16))  
p3
```

The DAPC shows a clear separation between healthy and disease tissue as well as between mTLE versus the mTORopathies. However, no clear separation could be achieved for the mTORopathies (FCD2a, FCD2b, TSC).

## mTORopathies

```{r}
grp2 <- amc_samples %>% 
    filter(disease %in% c("Focal cortical dysplasia type 2a",
                        "Focal cortical dysplasia type 2b",
                        "Tuberous Sclerosis Complex")) %>% 
  mutate(disease = case_when(disease == "Focal cortical dysplasia type 2a" ~ "FCD IIa",
                          disease == "Focal cortical dysplasia type 2b" ~ "FCD IIb",
                          disease == "Tuberous Sclerosis Complex" ~ "TSC",
                          disease == "Hippocampal Sclerosis" ~ "TLE-HS",
                          disease == "Control_Cortex" ~ "Cortex",
                          TRUE ~ "Hippocampus")) %>% 

  pull(disease) %>% as.factor() %>% as.numeric()

names(grp2) <- amc_samples %>% 
  filter(disease %in% c("Focal cortical dysplasia type 2a",
                        "Focal cortical dysplasia type 2b",
                        "Tuberous Sclerosis Complex")) %>% 
  pull(sample)


grp_name <- amc_samples %>% 
  filter(disease %in% c("Focal cortical dysplasia type 2a",
                        "Focal cortical dysplasia type 2b",
                        "Tuberous Sclerosis Complex")) %>% 
  mutate(disease = case_when(disease == "Focal cortical dysplasia type 2a" ~ "FCD IIa",
                          disease == "Focal cortical dysplasia type 2b" ~ "FCD IIb",
                          disease == "Tuberous Sclerosis Complex" ~ "TSC")) %>% 
  pull(disease) %>% as.factor() 
grp_name <- tibble(name = levels(grp_name),
                   value = 1:3)

sd <- d[,names(grp2)]
dapc2 <- dapc(t(sd), grp = grp2, n.pca = 8, n.da = 2)

summary(dapc2)

assigment <- tibble(prior = dapc2$grp,
                     posterior = dapc2$assign) %>% 
  mutate(prior_name = grp_name$name[match(prior, grp_name$value)],
         post_name = grp_name$name[match(posterior,
                                         grp_name$value)])
table(assigment$prior_name, assigment$post_name)

h <- table(assigment$prior_name, assigment$post_name)
h <- melt(h) %>% 
  as_tibble() %>% 
  mutate(Var2 = factor(gsub("mTLE", "TLE-HS", Var2),
                       levels = c("Hippocampus", "TLE-HS", 
                                  "Cortex", "FCD IIa", "FCD IIb", "TSC")),
         Var1 = factor(gsub("mTLE", "TLE-HS", Var1),
                       levels = c("Hippocampus", "TLE-HS", 
                                  "Cortex", "FCD IIa", "FCD IIb", "TSC")))

p4 <- ggplot(h, aes(x = Var1, y = Var2, fill = log10(value), label = value)) +
  geom_tile( color = "grey") +
  geom_text() +
  scale_fill_gradient2(na.value = "white") +
  labs(x = "Prior assignment", y = "Posterior assignment") + 
  theme_minimal() +
  theme(legend.pos = "none", legend.text = element_text(size = 16),
         text = element_text(size=14),
        axis.text=element_text(size=20), axis.title = element_text(size=18)) 
p4
ggsave(here("Data/12_PLSDA_DAPC/reassignment_mTORopathies.png"))


save(dapc2, 
     file = "/home/lfrancois/Documents/projects/AMC_collaboration_2019/Data/12_PLSDA_DAPC/dapc_mtor_clusters")
  
toPlot <- dapc2$ind.coord %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "sample") %>% 
  inner_join(amc_samples %>% select(sample, Cohort, 
                                    `tissue generic origin`, age),
             by = c("sample")) %>% 
  mutate(grp = dapc2$grp[sample]) %>% 
  as_tibble()
```

```{r}
cols <- control_palettes(palette = 'Categorical_12_colors')[c(1, 2, 3),] %>%
  pull(hex)
p5 <- ggplot(toPlot, aes(x = LD1, y = LD2, col = Cohort)) + 
  geom_point(size = 3, alpha = 0.7) +
  stat_ellipse(level=0.95) +
  # scale_color_brewer(palette = "Set2") +
  scale_color_manual(values=cols) +
  theme_bw() +
  theme(legend.pos = "none", text = element_text(size = 16), 
        axis.title = element_text(size = 18),
        axis.text=element_text(size=16)) +
  xlab("LD1") +
  ylab("LD2")   
p5
```

```{r}
library(ggpubr)

ggarrange(p1, labels = c("A)"),
          ggarrange(p3, p5, labels = c("B)", "C)"), ncol = 2), 
          ggarrange(p2, p4, labels = c("D)", "E)"), ncol = 2), 
          nrow = 3, common.legend = TRUE, legend = "top") %>% 
  ggexport(filename = here("Data/12_PLSDA_DAPC/combined_plot_dapc_results.png"), 
           height = 1000, width = 1600)
```

